Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-10-04 Latex_files/Thesis/chapters/chapter 2/main.tex
updates 04-10-2024

The scattering force arises from the scattered field of the trapping 
beam pushing the target object away from the focal point. While some 
trapping arrangements can induce transverse scattering forces for 
simple spheres the scattering force is only significant in the direction
of propagation \cite{Capitanio2002} The equilibrium position is found 
when the gradient force far exceeds the both the scattering force, 
and the thermal energy of the particle.

\subsubsection{Harmonic Traps}
\label{sec:harmonic_traps}
When a sphere is located at the centre of a laser the gradient force
will try to recentre the sphere if it is displaced. If the restoring 
force is proportional to the sphere's displacement this is referred 
to as a harmonic trap. The ratio of the restoring force to the 
displacement is called the trap stiffness (denoted by $\kappa$), and 
is an important characteristic of the trap. Shown below is a typical 
force displacement curve for a single sphere in a focused Gaussian beam. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{sphere_force_curve.png}
	\caption{Example of a sphere's ($r=1\ \mu m$) force-displacement curve 
			while moving along the x-axis, where $x=0$ is the centre of the
			beam. The harmonic regime is highlight by the black linear fit.}
	\label{fig:harmonic_trap}
\end{figure}

The strength of the optical trap is directly proportional to both 
the laser power and the contrast in refractive indices between the 
medium and the particle. Fully describing and calculating the 
optical forces is dependent on the ratio between the particle size 
and the trapping wavelength. Depending on this ratio, the optical 
forces can be described differently.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ray-Optics Regime}

The Ray-Optics model is the simplest to understand, this theory models 
the laser as a collection of individual 'rays' that propagate and are 
refracted according to Snell's Law. Based on the change in direction 
momentum is transferred to the target particle; with rays closest to 
the centre of the beam having greater intensity than those rays at 
the very edge of the beam. Consider a particle struck by two rays in 
a Gaussian beam, one coming close to the centre, and the other ray 
coming from the edge. As each ray is refracted by the target sphere a 
force is imparted onto it, the total force imparted is given by:
\begin{align}
	F_i = Q_i\frac{\Delta n P_i}{c}
\end{align}

where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference
in refractive indices between the solution and the target particle,
and $P_i$ is the power of the individual ray. The net force can be 
subdivided into its gradient and scattering components, where the latter
directs the particle to the centre of the beam, and the latter acts on 
the particle in the direction of beam propagation. For a beam with a
Gaussian intensity distribution $P_i$ will fall off as you move from
the centre of the beam. The ray optics model is ideal when dealing 
with larger particles whose diameter far exceed the wavelength of light
being scattered. While it can be useful in predicting the forces 
experienced by said particle's it does not fully capture the behaviour
of light when considering interference between different rays.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Lornez-Mie Theory}

@@

Where $k$ is the wave number of the electromagnetic radiation 
($k = 2\pi/\lambda$). This combined with the constraints of 
Maxwell's equations leaves very few exact solutions apart from 
spherical or planar waves, both of which can be converted between 
with relative ease. For example, a plane wave electric field can 
be expanded into spherical harmonics and likewise any spherical 
wave can be described as a combination of plane waves offset from 
one another. For a single plane wave the incident, internal, and 
scattered fields are given as: 
\begin{align}
	\label{eq:incident}
	{\bf E}_{\rm inc}({\bf r})
	&=
    E_0 \sum_{nm}^\infty \left[
    a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
    + b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
  \\
  \label{eq:internal}
  {\bf E}_{\rm int}({\bf r})
  &=
    E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left[
    -id_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
    \right]
  \\
  \label{eq:scattered}
  {\bf E}_{\rm scat}({\bf r})
  &=
    E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left[
      -i p_{nm}{\bf N}_{nm}^{(2)}({\bf r})+q_{nm}{\bf M}_{nm}^{(2)}({\bf r})
    \right] 
    \\
	\mathbf{E}_{tot}({\bf r}) &= 
	\begin{cases}
		{\bf E}_{\rm inc}({\bf r}) + {\bf E}_{\rm scat}({\bf r}) & \text{if \textbf{r} outside}
		\\ 
		{\bf E}_{\rm int}({\bf r}) & \text{if \textbf{r} inside}
	\end{cases}
\end{align}

where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
are the expansion coefficients of each of the fields, and $M_{nm}$ 
and $N_{nm}$ are the vector spherical harmonics. The superscripts
denote whether the field is inward or outward propagating; so a 
superscript (1) is inward propagating, and (2) is outward propagating.  
For the incident field computing its expansion coefficients is possible 
via analytical methods and are completely dependent on the beam 
conditions imposed by the user. However, computing the expansion 
coefficients of the internal and scattered fields is far more 
tedious depending on the shape of the target and what properties 
of the scattered field are desired --- see sec.~\ref{sec:scattering}.  
From an optical trapping perspective the force exerted by a focused 
electric field can be found by computing the Maxwell stress tensor 
which only requires the total magnitude of the incident and scattered 
fields. An alternative method is to compute the momentum carried by 
the incident and scattered beams and taking the difference between 
to measure the momentum transferred to the particle. This project 
utilises the results from Farsund \cite{Farsund1996} but in the 
form used by Crichton \cite{Crichton2000THEMD}. As an example the 
linear momentum carried by any beam can be computed via:
\begin{equation}
	\label{eq:linear_momentum_in}
	\begin{split}
		\langle P_{z,\ in} \rangle
		&=
		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}] \right.
		\\ 
		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
		\\
		& \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}] \right)
	\end{split}
\end{equation}

@@

Therefore the total force imparted to the trapped particle is thus:
\begin{align}
	\boldsymbol{F_z} =\frac{dP}{dt} = P_{z,\ in} -P_{z,\ out}
\end{align}

Where $a_{nm}, b{nm}, p_{nm}, \& q_{nm}$ are the beam expansion 
coefficients used in \eqref{eq:incident} \& \eqref{eq:scattered}.
Similar expressions are shown in \ref{sec:simulations} for 
calculating the optical torque. Lorenz-Mie theory can be applied 
to describe the scattering from any particle regardless of size, 
though as the size of the target particle changes - relative to 
the light being scattered - the theory describing the scattered 
light is better approximated by alternative theories. For 
particles that are far larger than the wavelength of light the 
scattering is best described by the Ray-Optics theory. In the 
opposite case the target particle is better approximated as a 
single electric dipole as described by the Rayleigh scattering 
field.

@@

The Rayleigh approximation is the limiting approximation for 
describing a particles scattering in an electromagnetic field 
whose wavelength is several times greater than the particle's 
size. The underlying theory is that a dielectric sphere can be 
treated as a dipole while in the presence of the electromagnetic 
field. In which case the scattering force is given simply by the 
scattering of the induced dipole, and the gradient force is due 
to the Lorentz force \cite{Gordon1973}. The gradient forces in the 
principle Cartesian axis are described by Harada et al 
\cite{YasuhiroHarada1996} in MKS units as a restorative rectangular 
force field:
\begin{align}
  F_{grad,x}
  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,y}
  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,z}
  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2}
    \nonumber \times I(r)
  \\ 
  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
  \\
  \text{where:}
  \nonumber
  \\
	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
\end{align}

Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is 
the beam waist, and $a$ is the radius of the particle. The scattering force 
however, is dependent on the effective scattering cross-sectional area. 
\begin{align}
  F_{\rm scat}
  &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
  \text{where:} \nonumber \\
  C_{pr} &= \frac{8}{3} \pi (ka)^4 a^2 \left(\frac{m^2-1}{m^2+2}\right)^2
\end{align}

The Rayleigh regime allows for easy computation of the gradient and 
scattering forces due to the assumption that the particle is a point 
dipole. Unlike Lorenz-Mie scattering, there is no need to compute the 
scattered field (see \eqref{eq:scattered}) for an individual sphere. 
Higher complexity scattering problems can be simplified by subdividing 
the particle into discrete dipoles (see Sec~\ref{sec:scattering}), 
in which case the scattered field is now necessary. As the target 
particle gets larger this assumption fails to accurately describe 
the trapping force. For particle sizes close to the laser wavelength 
the trapping force is best described by Lorenz-Mie theory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scattering methods}
\label{sec:scattering}
There are several methods available to calculate the scattered field 
produced by a particle. While one can compute this by directly applying
Lorenz-Mie theory, describing the interactions between focused beams and 
complicated shapes can be computationally slow and memory intensive. As a
result, several alternative scattering approaches have been developed that
allow for faster computation times simplifying the scattering problem.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{T-matrix Method}
The $T$-matrix method was first developed by Peter Waterman with his
research into acoustic wave scattering \cite{Waterman1969}, this 
would later be extended to electromagnetic waves. Sometimes  
referred to as the extended boundary condition method (ECBM), 
the method replaces the scatterer with a series of surface currents
over the targets surface. These currents are chosen so that the
electromagnetic field outside is identical to the original problem 
\cite{Wriedt1998}. This choice of currents negates the need to 
compute the internal fields which reduces the scattering problem 
to a system of linear equations relating the incident beam coefficients 
to the scattered beam coefficients. The relationship between each can 
be summarised as:
\begin{align}
	\label{eq:t_matrix}
	\begin{pmatrix}
		q_{mn} \\
		p_{mn} 
	\end{pmatrix}
	= \bold{T} 
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
	= \begin{pmatrix}
		T_{11} \ T_{12} \\
		T_{21} \ T_{22}
	\end{pmatrix}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}

Where $q_{mn}$, and $p_{mn}$ are the scattering beam coefficients, 
and $a_{mn}$, and $b_{mn}$ are the coefficients of the incident beam. 
For this project we utilised an extension of the ECBM known as 
multi-sphere $T$-matrix method (\textit{mstm}). Developed by Mackowski
\cite{Mackowski2011} the computational code computes the scattered 
field from each sphere within the target cluster and the incident 
field itself. The scattering between spheres slowly converges to a 
final result but modern codes truncate the calculations to an upper 
number of iterations (denoted by $N_{max}$) to fit a desired accuracy 
and computational time.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discrete Dipole Approximation}
The discrete dipole approximation (DDA) is a general method that can 
be applied to the scattering from particles of arbitrary composition
and geometry. Developed by Purcell and Pennypacker \cite{Purcell1973}, 
the DDA method approximates the particle as being constructed of dipoles.
Each dipole interacts with both the incident field and the scattered fields
from every other dipole surrounding it. The resulting scattered field is 
identical to the scattered field produced by direct integration of 
Eq.~\eqref{eq:internal} throughout the full particle volume \cite{Goedecke1988}.
The integral form for the internal electric field inside a scatterer is 
given as:
\begin{align}
	\mathbf{E(r)} = \mathbf{E_{inc}(r)} + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
	\chi(\mathbf{r'})\mathbf{E(r')}
\end{align}

Where $\mathbf{\bar{G}}$ is the Greens dyadic function of free space, which
defines the impulse response between two separate dipoles; and $\chi$ is the 
susceptibility of the medium, which describes the degree of polarisation of
the medium in the presence of an electric field. With the internal field 
calculated, the scattered field can be computed, one of the primary advantages
of DDA over the T-matrix method is that the composition of the target can 
be changed freely. When comparing different computational scattering methods,
the ECBM was found to be better suited for simulating the scattering of symmetric
targets as the ECBM can use the target's symmetry to speed up calculations \cite{Wriedt1998}. But when dealing with inhomogeneous media DDA is 
more efficient compared to ECBM.

@@

The Einstein model of diffusion expands upon the Fickian model by considering
the collisions between individual molecules. If we consider a single particle 
suspended in a solution it will experience multiple collisions with other 
molecules \cite{Gillespie2012a} over a given time interval $\delta t$. The 
Einstein molecule allows us to consider how individual molecules behave while
suspended in a surrounding fluid. Where it begins to falter is when we consider
bringing $\delta t$ to $0$; in this scenario the molecule experiences not many
but only individual collisions. The Einstein model does not consider the inertia 
of the molecule in question and so for very short time frames the model implies 
that the molecule's velocity changes instantly after each collision \cite{Gillespie2012a,
Gillespie2012b}. Furthermore, the kinetic energy of each collision is not limited
by the thermal energy of the system, meaning that using the Einstein model to 
predict a particle's trajectory to an finite degree of accuracy implies that 
the particle has infinite kinetic energy \cite{Gillespie2012b}. This failure to 
describe motion over smaller time frames is addressed in the Langevin model by 
accounting for the fluid drag of the system, where any sudden change in velocity 
must result in a proportionally opposed drag force \cite{Gillespie2012c}.

@@

\subsection{Power Spectrum Density (PSD)}
The power spectral density (PSD) method is by far the most versatile 
method for observing the dynamics of any object within an optical trap, 
allowing for fast calibration times while also quickly filtering out 
typical noise sources. Taking the Fourier transform of a particle's 
trajectory yields:
\begin{align}
	\hat{x}(f) = \frac{(2D)^{1/2}\hat{\eta}}{2\pi(f_c-if)}
\end{align}

where $\hat{\eta}$ is the Fourier transform of the white noise, 
where the values are exponentially distributed as opposed to being 
normally distributed in the time domain \cite{BergSoerensen2004}, 
the correlation function is given as:
\begin{align}
	\left<\hat{\eta_k}\hat{\eta_l^*}\right> = t_{msr} \delta_{k,l} \rightarrow \left< \eta^4 \right> = 2t_{msr}
\end{align}

>>>>

The scattering force arises from the scattered field of the trapping 
beam pushing the target object away from the focal point. While some 
trapping arrangements can induce transverse scattering forces for 
simple spheres the scattering force is only significant in the direction
of propagation \cite{Capitanio2002} The equilibrium position is found 
when the gradient force far exceeds both the the scattering force, 
and the thermal energy of the particle.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Harmonic Traps}
\label{sec:harmonic_traps}
When a sphere is located at the centre of a laser the gradient force
will try to recentre the sphere if it is displaced. If the restoring 
force is proportional to the sphere's displacement this is referred 
to as a harmonic trap. The ratio of the restoring force to the 
displacement is called the trap stiffness (denoted by $\kappa$), and 
is an important characteristic of the trap. Shown below is a typical 
force displacement curve for a single sphere in a focused Gaussian beam. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=1.1\linewidth]{sphere_force_curve.png}
	\caption{Example of a sphere's ($r=1\ \mu m$) force-displacement curve 
			while moving along the x-axis, where $x=0$ is the centre of the
			beam. The harmonic regime is highlight by the black linear fit.}
	\label{fig:harmonic_trap}
\end{figure}

The strength of the optical trap is directly proportional to both 
the laser power and the contrast in refractive indices between the 
medium and the particle. Fully describing and calculating the 
optical forces is dependent on the ratio between the particle size 
($a$) and the trapping wavelength ($\lambda$). The scattering theory 
for computing the optical force is dependent on this ratio.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Lornez-Mie Theory}

@@

Where $k$ is the wave number of the electromagnetic radiation 
($k = 2\pi/\lambda$). This combined with the constraints of 
Maxwell's equations leaves very few exact solutions apart from 
spherical or planar waves. For a laser of arbitrary shape the 
electric field can be described as a summation of partial 
spherical waves multiplied by an expansion coefficient. Depending 
on the beam in question the expansion coefficients can be tailored
to match the desired beam shape. For a Gaussian beam interacting 
with a particle of volume $V$, the incident, internal, and 
scattered fields are given as \cite{RanhaNeves2019}: 
\begin{align}
	\label{eq:incident}
	{\bf E}_{\rm inc}({\bf r})
	&=
	E_0 \sum_{n}^\infty\sum_{m=-n}^n \left[
	a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
	+ b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
	\\
	\label{eq:internal}
	{\bf E}_{\rm int}({\bf r})
	&=
	E_0 \sum_{n}^\infty \sum_{m=-n}^{n}\left[
	d_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
	\right]
	\\
	\label{eq:scattered}
	{\bf E}_{\rm scat}({\bf r})
	&=
	E_0 \sum_{n}^\infty \sum_{m=-n}^n\left[
	p_{nm}{\bf N}_{nm}^{(3)}({\bf r})+q_{nm}{\bf M}_{nm}^{(3)}({\bf r})
	\right] 
\end{align}

where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
are the expansion coefficients of each of the fields, and ${\bf M}_
{nm}$ and ${\bf N}_{nm}$ are the magnetic and electric vector 
spherical harmonics respectively. The superscripts denote whether 
the field is inward [(1)] or outward propagating [(3)]. By linearity 
of Maxwell's equations, its possible to relate the expansion 
coefficients of the scattered field directly to those of the incident 
field. As such we can write the scattering coefficients in terms of 
their incident components and the target's 'T-matrix'.
\begin{align}
	\label{eq:t_matrix}
	\begin{pmatrix}
		q_{mn} \\
		p_{mn} 
	\end{pmatrix}
	= \bold{T} 
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
	= \begin{pmatrix}
		T_{11} & T_{12} \\
		T_{21} & T_{22}
	\end{pmatrix}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}

Where the matrix \textbf{T} accounts for the targets shape, and 
refractive index. The $T$-matrix method can be easily 
applied to any solution of Maxwell's equations, for example Eq.~\eqref{eq:scattered} can be instead be rewritten as:
\begin{align}
{\bf E}_{\rm scat}({\bf r})
&=
E_0 \sum_{n}^\infty \sum_{m=-n}^n 
\begin{pmatrix}
	T_{11} & T_{12} \\
	T_{21} & T_{22}
\end{pmatrix}
\begin{pmatrix}
	a_{mn} \\
	b_{mn}
\end{pmatrix}
\begin{pmatrix}
	{\bf N}_{mn}^{(3)} & {\bf M}_{mn}^{(3)}
\end{pmatrix}
\end{align}

In order to compute the optical force and torque applied to a 
given object one simply needs to know the momentum carried by 
the incident and scattered fields. So long as we know the 
expansion coefficients of the incident beam we can find the 
scattered beam coefficients via Eq.~\eqref{eq:t_matrix} to 
compute the force. This project utilises the results from 
Farsund \cite{Farsund1996} but in the form used by Crichton \cite{Crichton2000THEMD}. As an example the linear momentum 
carried by any beam can be computed via:
\begin{equation}
	\label{eq:linear_momentum_in}
	\begin{split}
		\langle P_{z,\ in} \rangle
		&=
		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}] \right.
		\\ 
		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
		\\
		& \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}] \right)
	\end{split}
\end{equation}

@@

Therefore the total force imparted to the trapped particle is thus:
\begin{align}
	\boldsymbol{F_z} =\frac{dP}{dt} = P_{z,\ in} -P_{z,\ out}
	\label{eq:Fz}
\end{align}

\noindent
Where $a_{nm}, b{nm}, p_{nm}, \& q_{nm}$ are the beam expansion 
coefficients used in \eqref{eq:incident} \& \eqref{eq:scattered}.
Similar expressions are shown in Sec.~\ref{sec:simulations} for 
calculating the optical torque. Lorenz-Mie theory can be applied 
to describe the scattering from any particle regardless of size, 
though as the size of the target particle changes - relative to 
the wavelength of light - the theory describing the electromagnetic 
fields are better easier to approximate by alternative theories. 
When $a \gg \lambda$ the optical force is best described by the 
Ray-Optics model. In the opposite case the target particle is 
better approximated as a single electric dipole as described by 
the Rayleigh approximation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ray-Optics Regime}

The Ray-Optics model is the simplest to understand, this theory models 
the electromagnetic field as a collection of individual 'rays' that 
propagate and are refracted according to Snell's Law. Based on the 
change in direction momentum is transferred to the target particle; 
with rays closest to the centre of the beam having greater intensity 
than those rays at the very edge of the beam. Consider a particle 
struck by two rays in a Gaussian beam, one coming close to the centre, 
and the other ray coming from the edge. As each ray is refracted by 
the target sphere a force is imparted onto it, the total force 
imparted is given by:
\begin{align}
	F_i = Q_i\frac{\Delta n P_i}{c}
\end{align}

\noindent 
where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference
in refractive indices between the solution and the target particle,
and $P_i$ is the power of the individual ray. The net force can be 
subdivided into its gradient and scattering components, where the latter
directs the particle to the centre of the beam, and the latter acts on 
the particle in the direction of beam propagation. For a beam with a
Gaussian intensity distribution $P_i$ will fall off as you move from
the centre of the beam. The ray optics model is ideal when dealing 
with larger particles whose diameter far exceed the wavelength of light
being scattered. While it can be useful in predicting the forces 
experienced by said particle's it does not fully capture the behaviour
of light when considering interference between different rays.

@@

The Rayleigh approximation is for describing a particle when 
$a \ll \lambda$. The underlying theory is that a dielectric sphere 
can be treated as a dipole while in the presence of the 
electromagnetic field. In which case the scattering force is 
given simply by the scattering of the induced dipole, and the 
gradient force is due to the Lorentz force \cite{Gordon1973}. 
The gradient forces in the principle Cartesian axis are 
described by Harada et al \cite{YasuhiroHarada1996} in MKS units 
as a restorative rectangular force field:
\begin{align}
  F_{grad,x}
  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \label{eq:Rayleigh_start}\\
  F_{grad,y}
  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,z}
  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{z}/w_0}{1+(2\tilde{z})^2}
    \nonumber \times I(r)
  \\ 
  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
  \\
  \label{eq:Rayleigh_end}
  \text{where:}
  \nonumber
  \\
	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
\end{align}

\noindent
Where $m$ is the relative refractive index ($n_1/n_2$), $n_2$ is 
the particle's refractive index, $\omega_0$ is the beam waist, 
$a$ is the radius of the particle, and $\tilde{x}, \tilde{y}, 
\tilde{z}$ are simply the $x, y, z$ coordinates but scaled by the 
beam radius. The scattering force however, is dependent on the 
effective scattering cross-sectional area. 
\begin{align}
  F_{\rm scat}
  &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
  \text{where:} \nonumber \\
  C_{pr} &= \frac{8}{3} \pi (ka)^4 a^2 \left(\frac{m^2-1}{m^2+2}\right)^2
\end{align}

There is a clear difference between Eq.~\eqref{eq:Rayleigh_start}-
\eqref{eq:Rayleigh_end} and Eq.~\eqref{eq:Fz}, namely that the 
Rayleigh regime is not concerned with the state of the scattered
field. This is because if the particle is approximated by a point
dipole we can ignore the momentum transfer calculation and simply
compute the Lorenz force for a dipole in an electric field. This is 
not to say that the scattered field is not influenced by the particle, 
the scattered field can be related to the incident field like so:
\begin{align}
	\label{eq:Rayleigh_scattering}
	\begin{pmatrix}
		{\bf E}^s_\perp \\
		{\bf E}^s_\parallel 
	\end{pmatrix}
	= \begin{pmatrix}
		S_{11} & 0 \\
		0 & S_{22}
	\end{pmatrix}
	\begin{pmatrix}
		{\bf E}^i_\perp \\
		{\bf E}^i_\parallel
	\end{pmatrix}
\end{align}

\noindent
Where $S_{11}$, and $S_{22}$ are elements of the amplitude scattering 
matrix, this is analogous to a single plane wave scattering off a 
sphere. The difference between Eq.~\eqref{eq:Rayleigh_scattering} and
Eq.~\eqref{eq:t_matrix} is that the $T$-matrix accounts for the 
scattering from multiple plane waves that make up the laser. Because the 
particle can be approximated by a point dipole the scattering field is 
linearly related to the scattered field. While not important for optical 
force calculation higher complexity scattering problems can be simplified 
by subdividing the particle into discrete dipoles (see 
Sec~\ref{sec:scattering}), in which case the scattered field is now 
significant. As the target particle gets larger this assumption fails 
to accurately describe the trapping force \cite{Li2021}. For particle 
sizes close to the laser wavelength the trapping force is best described 
by Lorenz-Mie theory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scattering methods}
\label{sec:scattering}
There are several methods available to calculate the scattered field 
produced by a particle. While one can compute this by directly applying
Lorenz-Mie theory, describing the interactions between focused beams and 
complicated shapes can be computationally slow and memory intensive. As a
result, several alternative scattering approaches have been developed that
allow for faster computation times simplifying the scattering problem.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{T-matrix Method}

The $T$-matrix method was first developed by Peter Waterman with 
his research into acoustic wave scattering \cite{Waterman1969}, 
this would later be extended to electromagnetic waves. Sometimes 
referred to as the extended boundary condition method (ECBM), the 
method replaces the scatterer with a series of surface currents 
over the targets surface. These currents are chosen so that the 
electromagnetic field outside is identical to the original problem \cite{Wriedt1998}. By expanding the fields one can derive the 
individual elements of \textbf{T} until a desired accuracy is 
achieved.

For this project we utilised an extension of the ECBM known as 
multi-sphere $T$-matrix method (\textit{mstm}). Developed by 
Mackowski \cite{Mackowski2011} the computational code computes 
the scattered field from each sphere within the target cluster 
and the incident field itself. The scattering between spheres 
will converge to a final result but modern codes truncate 
the calculations to an upper number of iterations (denoted by 
$N_{max}$) to fit a desired accuracy and computational time. 
By centring the expansion at the origin of each sphere the 
individual sphere $T$-matrices can be computed, then using the
addition theorem, the target centred $T$-matrix is computed.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discrete Dipole Approximation}
The discrete dipole approximation (DDA) is a general method that 
can be applied to the scattering from particles of arbitrary 
composition and geometry. Developed by Purcell and Pennypacker \cite{Purcell1973}, the DDA method approximates the particle as 
being constructed of discrete dipoles. Each dipole interacts with 
both the incident field and the scattered fields from every other 
dipole surrounding it. The resulting scattered field is identical 
to the scattered field produced by direct integration of Eq.~\eqref{eq:internal} throughout the full particle volume 
\cite{Goedecke1988}. The integral form for the total electric field 
inside a scatterer is given as \cite{Wriedt1998}:
\begin{align}
	\mathbf{E(r)} = \mathbf{E}_{inc}(r) + \sum^{N_d}_{i}\mathbf{E}_{scat,i} =  \mathbf{E}_{inc}(r) + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
	\chi(\mathbf{r'})\mathbf{E(r')}
	\label{eq:DDA}
\end{align}

Where $\mathbf{\bar{G}}$ is the Greens dyadic function of free space, 
which defines the impulse response between two separate dipoles; and 
$\chi$ is the susceptibility of the medium, which describes the 
degree of polarisation of the medium in the presence of an electric 
field. One of the primary advantages of DDA over the T-matrix method 
is that the composition of the target can be changed freely. When 
comparing different computational scattering methods, the ECBM was 
found to be better suited for simulating the scattering of symmetric
targets as the ECBM can use the target's symmetry to speed up 
calculations \cite{Wriedt1998}. But when dealing with inhomogeneous 
media DDA is more efficient compared to ECBM.

The DDA method can be used to derive the target's $T$-matrix. After
generating the far field expansion of the electric field (Eq.~
\eqref{eq:DDA}), the vector spherical expansion far field 
(Eq.~\eqref{eq:scattered}) can be matched to it. Solving for the 
expansion coefficients $p_{nm}$, and $q_{nm}$ allows one to construct
the T-matrix of the target.

@@

The Einstein model of diffusion expands upon the Fickian model by 
considering the collisions between individual molecules. If we 
consider a single particle suspended in a solution it will experience 
multiple collisions with other molecules \cite{Gillespie2012a} over a 
given time interval $\delta t$. The Einstein molecule allows us to 
consider how individual molecules behave while suspended in a 
surrounding fluid. Where it begins to falter is when we consider 
bringing $\delta t$ to $0$; in this scenario the molecule experiences 
not many but only individual collisions. The Einstein model does not 
consider the inertia of the molecule in question and so for very short 
time frames the model implies that the molecule's velocity changes 
instantly after each collision \cite{Gillespie2012a,Gillespie2012b}. 
Furthermore, the kinetic energy of each collision is not limited by 
the thermal energy of the system, meaning that using the Einstein model 
to predict a particle's trajectory to an finite degree of accuracy 
implies that the particle has infinite kinetic energy \cite{Gillespie2012b}. 
This failure to describe motion over smaller time frames is addressed 
in the Langevin model by accounting for the fluid drag of the system, 
where any sudden change in velocity must result in a proportionally 
opposed drag force \cite{Gillespie2012c}.

@@

\subsection{Power Spectrum Density (PSD)}
The power spectral density (PSD) method is by far the most versatile 
method for observing the dynamics of any object within an optical trap, 
allowing for fast calibration times while also quickly filtering out 
typical noise sources. After recording a trajectory over some time 
$t_{msr}$, taking the Fourier transform of a particle's trajectory 
yields:
\begin{align}
	\hat{x}(f) = \frac{(2D)^{1/2}\hat{\eta}}{2\pi(f_c-if)}
\end{align}
